Something about cetacean heart rates and ballistocardiograms

### 2021-08-31 analysis/paper/paper.Rmd
Split Corky BCG figure into two

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)
library(cetaceanbcg)
library(tidyverse)
```

@@

```{r process-acc}
acc_400hz <- readRDS(here::here("analysis/data/derived_data/acc_400hz.rds"))
fs <- 400
fny <- fs / 2
acc_filter <- signal::butter(4, c(1.0, 25.0) / fny, type = "pass") 
apply_filter <- function(x) signal::filtfilt(acc_filter, x)
acc_400hz <- acc_400hz %>% 
  mutate(across(c(surge, sway, heave), apply_filter, .names = "{.col}_filt"))

# This window (in seconds) will be used for smoothing multiple signals and in
# heartbeat detection
window <- 0.25
prh_10hz <- readRDS(here::here("analysis/data/derived_data/prh_10hz.rds")) %>% 
  mutate(jerk_smooth = tma(jerk, window * 10))
```

```{r align-ecg}
ecg_beats <- readRDS(here::here("analysis/data/derived_data/ecg_beats.rds"))

acc_400hz$ecg_beat <- FALSE
beat_idx <- approx(acc_400hz$dt, 
                   seq_along(acc_400hz$dt), 
                   xout = ecg_beats$dt,
                   method = "constant")$y %>% 
  na.omit()
acc_400hz$ecg_beat[beat_idx] = TRUE

@@

```{r corky-bcg, fig.cap="A: The BCG signal is clearest in surge (longitudinal axis acceleration). Solid line: surge bandpass filtered to [1.0 25.0] Hz. Dashed line: local range of surge in 0.25 s sliding window (smoothed). Blue dots indicate peak of the BCG I-wave. B: Heart rates estimated from BCG (blue) match those from ECG (red). Temporal offset is due to both physiology and sensor clock differences. C: BCG signal is not as strong in heave (dorso-ventral axis) or sway (lateral axis) as in surge. The sway signal is likely due to movement of the chest cavity (see video S1). D: Some, but not all, heart beats are visible as peaks in the norm of the jerk vector."}
ca_tz <- "Etc/GMT+7"
bcg_begin <- lubridate::ymd_hms("2021-08-16 10:09:24.5", tz = ca_tz)
bcg_end <- lubridate::ymd_hms("2021-08-16 10:09:39", tz = ca_tz)
bcg_sample <- filter(acc_400hz, between(dt, bcg_begin, bcg_end))
bcg_sample_10hz <- filter(prh_10hz, between(dt, bcg_begin, bcg_end))

t <- theme_minimal() +
  theme(axis.title.x = element_blank(),
        panel.grid.minor = element_blank())

p2 <- ggplot(filter(bcg_sample, bcg_beat), aes(dt, bcg_bpm)) +
  geom_line() +
  geom_point(color = "blue") +
  geom_line(aes(y = ecg_bpm), filter(bcg_sample, ecg_beat)) +
  geom_point(aes(y = ecg_bpm), filter(bcg_sample, ecg_beat), color = "red") +
  scale_x_datetime(date_labels = "%H:%M:%S") +
  labs(y = "Heart rate (bpm)") +
  expand_limits(x = range(bcg_sample$dt), y = c(30, 70)) +
  t

cowplot::plot_grid(p1, p2, p3, p4, 
                   align = "v", 
                   axis = "lr",
                   nrow = 4, 
                   rel_heights = c(1.5, 0.75, 1, 1),
                   labels = "AUTO")
```

@@

bw_elg <- readRDS(here::here("analysis/data/derived_data/bw180905_53_elg.rds"))

bw_400hz <- readRDS(here::here("analysis/data/derived_data/bw180905_53_400hz.rds")) %>% 
  mutate(
    across(c(surge, sway, heave), apply_filter, .names = "{.col}_filt"),
    rid_left = approx(bw_elg$start, bw_elg$region_id, dt, "constant")$y,
    rid_right = approx(bw_elg$stop, bw_elg$region_id, dt, "constant", yleft = 0)$y + 1,
    region_id = ifelse(rid_left == rid_right, rid_left, NA)
  ) %>% 
  select(-c(rid_left, rid_right))

bw_10hz <- readRDS(here::here("analysis/data/derived_data/bw180905_53_10hz.rds")) %>% 
  mutate(jerk_smooth = tma(jerk, window * 10))
```

@@

foo <- lapply(unique(na.omit(bw_400hz$region_id)), bw_bcg_plot)
```

Figure \@ref(fig:corky-bcg) shows how we can have a caption and cross-reference for a plot.

Here is an example of inline code `r x` in the middle of a sentence.

>>>>

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  fig.height = 6,
  fig.width = 6,
  dpi = 300
)
library(cetaceanbcg)
library(tidyverse)
```

@@

```{r process-acc}
acc_400hz <- corky_400hz
fs <- 400
fny <- fs / 2
acc_filter <- signal::butter(4, c(1.0, 25.0) / fny, type = "pass") 
apply_filter <- function(x) signal::filtfilt(acc_filter, x)
acc_400hz <- acc_400hz %>% 
  mutate(across(c(surge, sway, heave), apply_filter, .names = "{.col}_filt"))

# This window (in seconds) will be used for smoothing multiple signals and in
# heartbeat detection
window <- 0.25
prh_10hz <- corky_10hz %>% 
  mutate(jerk_smooth = tma(jerk, window * 10))
```

```{r align-ecg}
acc_400hz$ecg_beat <- FALSE
beat_idx <- approx(acc_400hz$dt, 
                   seq_along(acc_400hz$dt), 
                   xout = ecg_beats$dt,
                   method = "constant")$y %>% 
  na.omit()
acc_400hz$ecg_beat[beat_idx] = TRUE

@@

```{r bcg-figs}
ca_tz <- "Etc/GMT+7"
bcg_begin <- lubridate::ymd_hms("2021-08-16 10:09:24.5", tz = ca_tz)
bcg_end <- lubridate::ymd_hms("2021-08-16 10:09:39", tz = ca_tz)
bcg_sample <- filter(acc_400hz, between(dt, bcg_begin, bcg_end))
bcg_sample_10hz <- filter(prh_10hz, between(dt, bcg_begin, bcg_end))

t <- theme_minimal() +
  theme(axis.title.x = element_blank(),
        panel.grid.minor = element_blank())
```

```{r corky-bcg-ecg, fig.cap="A: BCG, B: ECG, C: BPM comparison"}
# A: Surge + LSR
bcg_surge_lsr <- bcg_sample %>% 
  rename(Surge = surge_filt, LSR = surge_range) %>% 
  pivot_longer(c(Surge, LSR), 
               names_to = "var",
               values_to = "val")

pA <- ggplot(bcg_surge_lsr, aes(dt, val, linetype = var)) +
  geom_line(size = 0.2) +
  geom_point(data = filter(bcg_surge_lsr, bcg_beat), color = "blue") +
  scale_x_datetime(date_labels = "%H:%M:%S") +
  scale_linetype_manual(values = c(Surge = "solid", LSR = "dashed")) +
  labs(y = bquote(Surge~(m~s^{-2}))) +
  expand_limits(x = max(bcg_sample$dt) + 1) +
  t +
  theme(legend.position = "none")

# B: ECG
ecg_sample <- corky_ecg %>% 
  filter(between(dt, first(bcg_sample$dt), last(bcg_sample$dt)))
pB <- ggplot(ecg_sample, aes(dt, ecg_uV)) +
  geom_line(size = 0.2) +
  geom_point(data = filter(ecg_sample, ecg_beat), color = "red") +
  scale_x_datetime(date_labels = "%H:%M:%S") +
  labs(y = bquote(ECG~(mu * V))) +
  expand_limits(x = max(bcg_sample$dt) + 1) +
  t

# C: BCG + ECG
pC <- ggplot(filter(bcg_sample, bcg_beat), aes(dt, bcg_bpm)) +
  geom_line(size = 0.2) +
  geom_line(aes(y = ecg_bpm), filter(bcg_sample, ecg_beat)) +
  geom_point(color = "blue") +
  geom_point(aes(y = ecg_bpm), filter(bcg_sample, ecg_beat), color = "red") +
  scale_x_datetime(date_labels = "%H:%M:%S") +
  labs(y = "Heart rate (bpm)") +
  expand_limits(x = range(bcg_sample$dt) + c(0, 1), y = c(30, 70)) +
  t

cowplot::plot_grid(pA, pB, pC,
                   align = "v", 
                   axis = "lr",
                   nrow = 3, 
                   rel_heights = c(1.5, 1.0, 0.5),
                   labels = "AUTO")
```

```{r corky-bcg-acc, fig.cap="A: The BCG signal is clearest in surge (longitudinal axis acceleration). Solid line: surge bandpass filtered to [1.0 25.0] Hz. Dashed line: local range of surge in 0.25 s sliding window (smoothed). Blue dots indicate peak of the BCG I-wave. B: BCG signal is not as strong in heave (dorso-ventral axis) or sway (lateral axis) as in surge. The sway signal is likely due to movement of the chest cavity (see video S1). C: Some, but not all, heart beats are visible as peaks in the norm of the jerk vector."}
# A: Surge + LSR
# pA as above

# B: Sway, heave
bcg_sway_heave <- bcg_sample %>% 
  rename(Sway = sway_filt, Heave = heave_filt) %>% 
  pivot_longer(c(Sway, Heave), 
               names_to = "axis",
               values_to = "acceleration")
pB <- ggplot(bcg_sway_heave, aes(dt, acceleration)) +
  geom_line(size = 0.2) +
  geom_point(data = filter(bcg_sway_heave, bcg_beat), color = "blue") +
  scale_x_datetime(date_labels = "%H:%M:%S") +
  facet_grid(axis ~ .) +
  labs(y = bquote(Acc~(m~s^{-2}))) +
  t

# C: Jerk
pC <- ggplot(bcg_sample_10hz, aes(dt, jerk_smooth)) +
  geom_line(size = 0.2) +
  geom_point(data = filter(bcg_sample_10hz, bcg_beat), color = "blue") +
  scale_x_datetime(date_labels = "%H:%M:%S") +
  labs(y = bquote(Jerk~(m~s^{-3}))) +
  expand_limits(y = 0) +
  t

cowplot::plot_grid(pA, pB, pC,
                   align = "v", 
                   axis = "lr",
                   nrow = 3, 
                   # rel_heights = c(1, 1, ),
                   labels = "AUTO")
```

@@

bw_elg <- bw180905_53_elg

bw_400hz <- bw180905_53_400hz %>% 
  mutate(
    across(c(surge, sway, heave), apply_filter, .names = "{.col}_filt"),
    rid_left = approx(bw_elg$start, bw_elg$region_id, dt, "constant")$y,
    rid_right = approx(bw_elg$stop, bw_elg$region_id, dt, "constant", yleft = 0)$y + 1,
    region_id = ifelse(rid_left == rid_right, rid_left, NA)
  ) %>% 
  select(-c(rid_left, rid_right))

bw_10hz <- bw180905_53_10hz %>% 
  mutate(jerk = njerk(aw, 10),
         jerk_smooth = tma(jerk, window * 10))
```

@@

# foo <- lapply(unique(na.omit(bw_400hz$region_id)), bw_bcg_plot)
```

Figure \@ref(fig:corky-bcg-ecg) shows how we can have a caption and cross-reference for a plot.
